---
title: "Geopoll pretest demographic data analysis"
author: "Brian"
date: "11/27/2020"
output: html_document
---

```{r, libraries}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)
```

## exploring the dataset
```{r, load data}
geopoll_df<-read_csv("./www/geopollpretest.csv")
head(geopoll_df)
names(geopoll_df)
str(geopoll_df)
summary(geopoll_df)
```

## what is the age distribution of the respondents?
## Insights
### 1. The distribution is normal.
### 2. Majority of the respondents are between 25 and 31 years of age.
### N/B add a summary(geopoll_df$age) table to display the exact values.
```{r, age distribution}
ggplot(data = geopoll_df,aes(y=Age))+
  geom_boxplot(aes(fill="red"))+
  labs(
    title = "Distribution of geopoll respondents by age."
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )

ggsave("./www/age_distribution.png")
```

## what is the age group distribution of the respondents?
```{r, age group percentage distribution}
geopoll_agegroup<-geopoll_df %>% count(AgeGroup) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_agegroup)
ggplot(data = geopoll_agegroup,aes(x=reorder(AgeGroup,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by age group.",
    y="Percentage of respondents",
    x="Age Group"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/agegroup_distribution.png")
```

## what is the gender percentage distribution of the respondents?
```{r}
geopoll_gender<-geopoll_df %>% count(Gender) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_gender)
ggplot(data = geopoll_gender,aes(x=Gender,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by gender.",
    y="Percentage of respondents",
    x="Gender"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/gender_distribution.png")
```


## what is the percentage distribution of geopoll respondents by region?
```{r}
geopoll_region<-geopoll_df %>% count(Region) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_region)
ggplot(data = geopoll_region,aes(x=reorder(Region,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by region.",
    y="Percentage of respondents",
    x="Region"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/region_distribution.png")
```


## what is the percentage distribution of geopoll respondents by district?
```{r}
geopoll_district<-geopoll_df %>% count(District) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_district)

ggplot(data = geopoll_district,aes(x=reorder(District,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by district.",
    y="Percentage of respondents",
    x="District"
  )+
  guides(
    fill=FALSE
  )+
  coord_flip()+
  theme(
    axis.text.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/district_distribution.png")
```


## what is the percentage distribution of geopoll respondents by area
```{r}
geopoll_area<-geopoll_df %>% count(Urban_Rural) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_area)

ggplot(data = geopoll_area,aes(x=Urban_Rural,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by area type.",
    y="Percentage of respondents",
    x="Area Type"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/areatype_distribution.png")
```

## ACCESS TO WATER SUPPLY.
### In the past month, how did you get your drinking water?
```{r}
# geopoll_watersupply<-geopoll_df %>% select(11:18,24:31) %>% gather(key = "Watersupply",value = "status",WaterSupply_Tap:'WaterSupplyBefore_Spring/River')%>% filter(status=="Yes")%>% count(Watersupply,status) %>% mutate(percentage_people=round(100*n/sum(n),0))
geopoll_watersupply<-read_csv("./www/watersupply.csv")
#head(geopoll_watersupply)
ggplot(data = geopoll_watersupply,aes(x=Watersupply,y=percentage_people))+
  geom_col(aes(fill=covid),position = "dodge")+
  scale_fill_discrete(labels=c("Before COVID19","During COVID19"))+
  labs(
    title = "Drinking water supply.",
    y="Percentage of respondents",
    x="Water source"
  )+
  theme(
    axis.text.x = element_text(size = 8,angle = 45,margin = margin(t=15)),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14),
    legend.title = element_blank()
  )

ggsave("./www/drinkingwater_supply.png")
```


## Where is the water source that you use?
```{r}
geopoll_watersource<-geopoll_df %>% count(SourceLocation) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_watersource)

ggplot(data = geopoll_watersource,aes(x=SourceLocation,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by water source location.",
    y="Percentage of respondents",
    x="Water source"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/watersource_distribution.png")
```

## In the past 1 week, how many days was the piped water available?
```{r}
#filtering before and after flow days
geopoll_flowdays<-geopoll_df %>% select(10,19,32)
geopoll_flowdays2<-gather(geopoll_flowdays,key = "Type",value ="Days",FlowDays,FlowDaysBefore)
#head(geopoll_flowdays2)
ggplot(data = geopoll_flowdays2,aes(y=Days))+
  geom_boxplot(aes(fill=Type))+
  scale_fill_discrete(labels=c("FlowDays during COVID","FlowDays before COVID"))+
  facet_wrap(~Urban_Rural)+
  labs(
    title = "FlowDays distribution per area type"
  )+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14),
    legend.title = element_blank()
  )
ggsave("./www/flowdays_distribution.png")
```

## On the days when piped water is available, how many hours per day, on average does it flow?

```{r}
#filtering before and after flow hours
geopoll_flowhrs<-geopoll_df %>% select(10,20,33)
geopoll_flowhrs2<-gather(geopoll_flowhrs,key = "Type",value = "Hours",FlowHours,FlowHoursBefore)
#head(geopoll_flowhrs2)
ggplot(data = geopoll_flowhrs2,aes(y=Hours))+
  geom_boxplot(aes(fill=Type))+
  scale_fill_discrete(labels=c("FlowHours during COVID","FlowHours before COVID"))+
  facet_wrap(~Urban_Rural)+
  labs(
    title = "FlowHours distribution per area type"
  )+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14),
    legend.title = element_blank()
  )
ggsave("./www/flowhours_distribution.png")
```


## how much did you pay for drinking water in the past month?
```{r}
#filtering before and after payments
geopoll_pay<-geopoll_df %>% select(10,22,34) %>% gather(key = "Type",value = "Pay",WaterPay,WaterPayBefore)
#head(geopoll_pay)
ggplot(data = geopoll_pay,aes(y=Pay))+
  geom_boxplot(aes(fill=Type))+
  scale_fill_discrete(labels=c("Payment during COVID","Payment before COVID"))+
  facet_wrap(~Urban_Rural)+
  labs(
    title = "Water payment per area type",
    y="GHC"
  )+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14),
    legend.title = element_blank()
  )
ggsave("./www/water_payment.png")
```

