---
title: "Geopoll pretest data analysis"
author: "Brian"
date: "11/27/2020"
output: html_document
---

```{r, libraries}
library(dplyr)
library(ggplot2)
library(readr)
```

## exploring the dataset
```{r, load data}
geopoll_df<-read_csv("./www/geopollpretest.csv")
head(geopoll_df)
names(geopoll_df)
str(geopoll_df)
summary(geopoll_df)
```

## what is the age distribution of the respondents?
## Insights
### 1. The distribution is normal.
### 2. Majority of the respondents are between 25 and 31 years of age.
### N/B add a summary(geopoll_df$age) table to display the exact values.
```{r, age distribution}
ggplot(data = geopoll_df,aes(y=Age))+
  geom_boxplot(aes(fill="red"))+
  labs(
    title = "Distribution of geopoll respondents by age."
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )

ggsave("./www/age_distribution.png")
```

## what is the age group distribution of the respondents?
```{r, age group percentage distribution}
geopoll_agegroup<-geopoll_df %>% count(AgeGroup) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_agegroup)
ggplot(data = geopoll_agegroup,aes(x=AgeGroup,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by age group.",
    y="Percentage of respondents",
    x="Age Group"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/agegroup_distribution.png")
```

## what is the gender percentage distribution of the respondents?
```{r}
geopoll_gender<-geopoll_df %>% count(Gender) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_gender)
ggplot(data = geopoll_gender,aes(x=Gender,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by gender.",
    y="Percentage of respondents",
    x="Gender"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/gender_distribution.png")
```


## what is the percentage distribution of geopoll respondents by region?
```{r}
geopoll_region<-geopoll_df %>% count(Region) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_region)
ggplot(data = geopoll_region,aes(x=reorder(Region,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by region.",
    y="Percentage of respondents",
    x="Region"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/region_distribution.png")
```


## what is the percentage distribution of geopoll respondents by district?
```{r}
geopoll_district<-geopoll_df %>% count(District) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_district)

ggplot(data = geopoll_district,aes(x=reorder(District,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by district.",
    y="Percentage of respondents",
    x="District"
  )+
  guides(
    fill=FALSE
  )+
  coord_flip()+
  theme(
    axis.text.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/district_distribution.png")
```


## what is the percentage distribution of geopoll respondents by area
```{r}
geopoll_area<-geopoll_df %>% count(Urban_Rural) %>% mutate(percentage_people=round(100*n/sum(n),0))
#head(geopoll_area)

ggplot(data = geopoll_area,aes(x=Urban_Rural,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by area type.",
    y="Percentage of respondents",
    x="Area Type"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.y = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14)
  )
ggsave("./www/areatype_distribution.png")
```

