---
title: "Ghana Geopoll pretest data analysis."
output: html_document
---

## **Introduction**
### The report displays analysis of Ghana Geopoll pretest survey.
### The survey results has two sections:
*  ### Demographic analysis.
*  ### Water access analysis. 

```{r,libraries, include=FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(tidyr)

```


```{r, load data, include=FALSE}
geopoll_df<-read_csv("./www/geopollpretest.csv")
```

## 1. **Demographic analysis**
### What is the age distribution of the respondents?
```{r, age_distribution,echo=FALSE,fig.height=6,fig.width=14,fig.align='left'}
ggplot(data = geopoll_df,aes(y=Age))+
  geom_boxplot(aes(fill="red"))+
  labs(
    title = "Distribution of geopoll respondents by age."
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    plot.title = element_text(size = 18)
  )

```

### What is the age group distribution of the respondents?
```{r,agegroup data treatment,include=FALSE}
geopoll_agegroup<-geopoll_df %>% count(AgeGroup) %>% mutate(percentage_people=round(100*n/sum(n),0))

```

```{r, agegroup_distribution,echo=FALSE,fig.height=6,fig.width=14,fig.align='left'}
ggplot(data = geopoll_agegroup,aes(x=reorder(AgeGroup,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by age group.",
    y="Percentage of respondents",
    x="Age Group"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 18)
  )
```

### What is the percentage distribution of the respondents by gender?
```{r, gender data treatment, include=FALSE}
geopoll_gender<-geopoll_df %>% count(Gender) %>% mutate(percentage_people=round(100*n/sum(n),0))
```

```{r, gender distribution,echo=FALSE,fig.height=6,fig.width=14,fig.align='left'}
ggplot(data = geopoll_gender,aes(x=Gender,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by gender.",
    y="Percentage of respondents",
    x="Gender"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 18)
  )

```

### What is the percentage distribution of geopoll respondents by region?
```{r, region data treatment,include=FALSE}
geopoll_region<-geopoll_df %>% count(Region) %>% mutate(percentage_people=round(100*n/sum(n),0))
```

```{r, region distribution,echo=FALSE,fig.height=6,fig.width=14,fig.align='left'}
ggplot(data = geopoll_region,aes(x=reorder(Region,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title ="Percentage distribution of geopoll respondents by region.",
    y="Percentage of respondents",
    x="Region"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 18)
  )

```

### What is the percentage distribution of geopoll respondents by district?
```{r, district data treatment, include=FALSE}
geopoll_district<-geopoll_df %>% count(District) %>% mutate(percentage_people=round(100*n/sum(n),0))
```

```{r, district distribution,echo=FALSE,fig.height=6,fig.width=14,fig.align='left'}
ggplot(data = geopoll_district,aes(x=reorder(District,percentage_people),y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title = "Percentage distribution of geopoll respondents by district.",
    y="Percentage of respondents",
    x="District"
  )+
  guides(
    fill=FALSE
  )+
  coord_flip()+
  theme(
    axis.text.x = element_text(size = 14),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 18)
  )

```

### What is the percentage distribution of geopoll respondents by area type?
```{r, area type data treatment, include=FALSE}
geopoll_area<-geopoll_df %>% count(Urban_Rural) %>% mutate(percentage_people=round(100*n/sum(n),0))
```

```{r, area type distribution, echo=FALSE,fig.height=6,fig.width=14,fig.align='left'}
ggplot(data = geopoll_area,aes(x=Urban_Rural,y=percentage_people))+
  geom_col(aes(fill="red"))+
  labs(
    title ="Percentage distribution of geopoll respondents by area type.",
    y="Percentage of respondents",
    x="Area Type"
  )+
  guides(
    fill=FALSE
  )+
  theme(
    axis.text.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.y = element_text(size = 14),
    plot.title = element_text(size = 18)
  )

```

